### 2019-03-16 micro4_2.tex
Micro

\begin{enumerate}
	\item The monopoly's problem is:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda)(x_2 - w_2)
	\end{align*}
	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that
	\begin{align*}
	w_1^* = c_1(x_1^*)\\
	w_2^* = c_2(x_2^*)
	\end{align*}
	Assuming interior solution FOC is:
	\begin{align*}
	\begin{cases}
	1 - c_1'(x_1^*) = 0\\
	1 - c_2'(x_2^*) = 0
	\end{cases}\\
	\end{align*}
	\begin{align*}
	c_1'(x_1^*) = c_2'(x_2^*) \to x_1^* > x_2^* \to c_1(x_1^*) > c_1(x_2^*) \to w_2^* - c_1(x_2^*) > w_1^* - c_1(x_1^*) = 0
	\end{align*}
	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:
	\begin{align*}
	\begin{cases}
	w_1^* - c_2(x_1^*) \le w_2^* - c_2(x_2^*)\\
	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
	\end{cases}\\
	\begin{cases}
	c_1(x_1^*) - c_2(x_1^*) \le 0\\
	c_2(x_2^*) - c_1(x_2^*) \le 0
	\end{cases}
	\end{align*}
	Since $c_2(x) > c_1(x) \forall x > 0$ it follows that $x_2^* = 0$.
\end{enumerate}
\end{document}

>>>>

\begin{enumerate}
	\item The monopoly's problem is:
	\begin{align*}
	\underset{(x_1, w_1), (x_2, w_2)}{\max}\ \lambda(x_1 - w_1) + (1 - \lambda)(x_2 - w_2)
	\end{align*}
	Since outside options of workers are 0, the optimal contracts monopoly proposes to workers should be such that
	\begin{align*}
	w_1^* = c_1(x_1^*)\\
	w_2^* = c_2(x_2^*)
	\end{align*}
	Assuming interior solution FOC is:
	\begin{align*}
	\begin{cases}
	1 - c_1'(x_1^*) = 0\\
	1 - c_2'(x_2^*) = 0
	\end{cases}\\
	\end{align*}
	\begin{align*}
	c_1'(x_1^*) = c_2'(x_2^*) \to x_1^* > x_2^* \to c_1(x_1^*) > c_1(x_2^*) \to w_2^* - c_1(x_2^*) > w_1^* - c_1(x_1^*) = 0
	\end{align*}
	That means that at any interior solution high type has incentive to deviate and mimic to low type. Hence we should add incentive compatibility constraints. The IC constraints are:
	\begin{align*}
	\begin{cases}
	w_1^* - c_2(x_1^*) \le w_2^* - c_2(x_2^*)\\
	w_2^* - c_1(x_2^*) \le w_1^* - c_1(x_1^*)
	\end{cases}\\
	\begin{cases}
	c_1(x_1^*) - c_2(x_1^*) \le 0\\
	c_2(x_2^*) - c_1(x_2^*) \le 0
	\end{cases}
	\end{align*}
	Since $c_2(x) > c_1(x)\ \forall\ x > 0$ ($g(x) = c_2(x) - c_1(x)$, $g(0) = 0, g'(x) = c'_2(x) - c_1'(x) > 0$ hence $g(x)$ is strictly increasing, since $g(0) = 0$ then $g(x)>0\ \forall\ x > 0$) it follows that $x_2^* = 0$. That means that monopoly should offer the following contracts:
	\begin{align*}
	&(0, 0)\ \text{ which is supposed to be for low type}\\
	&(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ which is supposed to be for high type}
	\end{align*}
	\item At the competitive equilibrium firms should earn 0 profit. That means that if firms can observe types they should offer the following contracts to both types of workers:
	\begin{align*}
	(x_1, x_1), (x_2, x_2)
	\end{align*}
	Since firms are competing for workers, at the equilibrium both firms should offer workers such contracts that would maximize their utility (otherwise, if any of firm offer another contract, the rival firm can deviate, offer maximizing utility contract, and hire all workers). Of course if firms offer maximizing utility contracts workers have no incentives to deviate and choose a contract which was supposed to be for another type. As a result the firms should propose contracts:
	\begin{align*}
	(c_1'^{-1}(1), c_1'^{-1}(1))\\
	(c_2'^{-1}(1), c_2'^{-1}(1))
	\end{align*}
	\item If monopoly can distinguish between workers, it should propose each type a contract which would maximize its profit and leave each type of workers with zero utility. That is, the contracts should be:
	\begin{align*}
	(w_1 = c_1(c_1'^{-1}(1)), x_1 = c_1'^{-1}(1))\ \text{ high type}\\
	(w_2 = c_2(c_2'^{-1}(1)), x_2 = c_2'^{-1}(1))\ \text{ low type}
	\end{align*}
	If competitive firms can observe types of workers they anyway have to propose such contracts which would maximize workers' utility because otherwise the rival firm hire all workers. That means that firms should propose the same contracts as in the case of unobservable types. That is, first best is achieved under competition and is not achieved under monopoly.
	\item \begin{itemize}
		\item Monopoly, observable types:
		\begin{align*}
		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		x_1^* = 1,\ x_2^* = \frac{1}{2}\\
		w_1^* = \frac{1}{2},\ w_2^* = \frac{1}{4}
		\end{align*}
		\item Monopoly, unobservable types:
		\begin{align*}
		\underset{x_1, x_2}{\max}\ \frac{1}{2}\left(x_1 -\frac{x_1^2}{2}\right) + \frac{1}{2} \left(x_2 - x_2^2\right)\\
		s.t.\ 	\begin{cases}
		\frac{x_1^2}{2} - x_1^2 \le 0\\
		x_2^2 - \frac{x_2^2}{2} \le 0
		\end{cases}\\
		x_2^* = 0, x_1^* = 1\\
		w_2 = 0, w_1 = \frac{1}{2}
		\end{align*}
		\item Competition, observable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
		\item Competition, unobservable types:
		\begin{align*}
		x_1^* = 1, w_1^* = 1\\
		x_2^* = \frac{1}{2}, w_2^* = \frac{1}{2}
		\end{align*}
	\end{itemize}
\end{enumerate}
\section*{Insurance}
Consider the competitive screening model of the insurance market with two types of
consumers $H$ and $L$. All consumers have initial wealth $W > 0$ and may face a loss $d > 0$,
$d < W$. The probability of facing the loss is $p_L$ for the $L$ type of the consumer, and
$p_H$ for the $H$ type, where $p_H > p_L$. Multiple competitive insurance firms offer insurance
contracts against the loss to either type of the consumer. The probabilities of loss and the
utility functions for both types of consumers are common knowledge to all participants.
For parts (1) and (2), assume that all consumers have the same strictly concave
(Bernoulli) utility of wealth $u(W)$ with $u'(W) > 0$ and $u''(W) < 0$. That is, the two types
of consumers are distinguished only by the difference between $p_L$ and $p_H$
\begin{enumerate}
	\item  Assume that the types of the consumers are publicly observable. Draw a graph
	representing the equilibrium outcome of this model. Explain. Derive the utility levels offered to either type of the consumer in this equilibrium.
	\item Now assume that the types of the consumers are only known to the consumers themselves. Draw a graph to represent the separating equilibrium of this model. Explain the reasoning behind your graph. Compare the outcome to (1) and comment on
	the source of difference/similarity.
	\item Now suppose the $H$ types are more risk averse - with utility function $v(W)$ instead
	of $u(W)$. As above, probabilities of loss and the utility functions for both types of
	consumers are common knowledge, but the types of the consumers are only known
	to the consumers themselves. Draw a new graph to show the effect of this change in
	utility function on the contracts in the separating equilibrium. Suggest an intuitive
	explanation for why this change from $u(W)$ to $v(W)$ increases or reduces the utility
	for good types in the separating equilibrium.
\end{enumerate}

\begin{enumerate}
	\item 
\end{enumerate}
\end{document}
